
\begin{poem}
{A Valediction Forbidden Mourning}
{John Donne}

AS virtuous men pass mildly away, 
And whisper to their souls to go, 
Whilst some of their sad friends do say, 
``Now his breath goes,'' and some say, ``No.''\\

So let us melt, and make no noise,
No tear-floods, nor sigh-tempests move; 
'Twere profanation of our joys 
To tell the laity our love.\\

Moving of th' earth brings harms and fears; 
Men reckon what it did, and meant; 
But trepidation of the spheres, 
Though greater far, is innocent.\\

Dull sublunary lovers' love 
-- Whose soul is sense -- cannot admit 
Of absence, 'cause it doth remove 
The thing which elemented it.\\

But we by a love so much refined, 
That ourselves know not what it is, 
Inter-assur\`{e}d of the mind, 
Care less, eyes, lips and hands to miss.\\

Our two souls therefore, which are one, 
Though I must go, endure not yet 
A breach, but an expansion, 
Like gold to aery thinness beat.\\

If they be two, they are two so 
As stiff twin compasses are two; 
Thy soul, the fix'd foot, makes no show 
To move, but doth, if th' other do.\\

And though it in the centre sit, 
Yet, when the other far doth roam, 
It leans, and hearkens after it, 
And grows erect, as that comes home.\\

Such wilt thou be to me, who must, 
Like th' other foot, obliquely run; 
Thy firmness makes my circle just, 
And makes me end where I begun.\\
\end{poem}

\begin{poem}
{The Good-Morrow}
{John Donne}

I wonder by my troth, what thou and I 
Did, till we lov'd? Were we not wean'd till then, 
But suck'd on country pleasures, childishly? 
Or snorted we in the seven sleepers' den? 
'Twas so; but this, all pleasures fancies be. 
If ever any beauty I did see, 
Which I desir'd, and got, 'twas but a dream of thee.\\

And now good morrow to our waking souls, 
Which watch not one another out of fear; 
For love, all love of other sights controls, 
And makes one little room, an everywhere. 
Let sea-discoverers to new worlds have gone, 
Let maps to other, worlds on worlds have shown, 
Let us possess one world, each hath one, and is one.\\

My face in thine eye, thine in mine appears, 
And true plain hearts do in the faces rest; 
Where can we find two better hemispheres, 
Without sharp north, without declining west? 
Whatever dies, was not mix'd equally; 
If our two loves be one, or, thou and I 
Love so alike, that none do slacken, none can die.\\
\end{poem}

\begin{poem}
{Only Our Love}
{John Donne}

Only our love hath no decay; 
This, no tomorrow hath, nor yesterday, 
Running it never runs from us away, 
But truly keeps his first, last, everlasting day.
\end{poem}

\begin{poem}
{Sweetest Love, I do not go}
{John Donne}
 
Sweetest love, I do not go 
For weariness of thee, 
Nor in hope the world can show 
A fitter love for me; 
But since that I 
Must die at last, 'tis best 
Thus to use myself in jest, 
By feigned death to die.\\

Yesternight the sun went hence, 
And yet is here to - day; 
He hath no desire nor sense, 
Nor half so short a way. 
Then fear not me, 
But believe that I shall make 
Hastier journeys, since I take 
More wings and spurs than he.\\

O how feeble is man's power, 
That, if good fortune fall, 
Cannot add another hour, 
Nor a lost hour recall. 
But come bad chance, 
And we join to it our strength, 
And we teach it art and length, 
Itself o'er us t' advance.\\

When thou sigh'st, thou sigh'st no wind, 
But sigh'st my soul away; 
When thou weep'st, unkindly kind, 
My life's blood doth decay. 
It cannot be 
That thou lov'st me as thou say'st, 
If in thine my life thou waste, 
That art the best of me.\\

Let not thy divining heart 
Forethink me any ill. 
Destiny may take thy part 
And may thy fears fulfil; 
But think that we 
Are but turned aside to sleep: 
They who one another keep 
Alive, ne'er parted be.\\
\end{poem}

\begin{poem}
{Loves' Infinitenesse}
{John Donne}
 
If yet I have not all thy love, 
Deare, I shall never have it all, 
I cannot breath one other sigh, to move; 
Nor can intreat one other teare to fall. 
And all my treasure, which should purchase thee, 
Sighs, teares and oathes, and letters I have spent, 
Yet no more can be due to mee, 
Then at the bargaine made was ment, 
If then thy gift of love were partiall, 
That some to mee, some should to others fall, 
   Deare, I shall never have Thee All.\\

Or if then thou gavest mee all, 
All was but All, which thou hadst then, 
But if in thy heart, since, there be or shall, 
New love created bee, by other men, 
Which have their stocks intire, and can in teares, 
In sighs, in oathes, and letters outbid mee, 
This new love may beget new feares, 
For, this love was not vowed by thee, 
And yet is was, thy gift being generall, 
The ground, thy heart is mine, what ever shall 
   Grow there, deare, I should have it all.\\

Yet I would not have all yet, 
Hee that hath all can have no more, 
And since my love doth every day admit 
New growth, thou shouldst have new rewards in store, 
Thou canst not every day give me thy heart, 
If thou canst give it, then thou never gavest it: 
Loves riddles are, that though thy heart depart, 
It stayes at home, and thou with losing savest it: 
But wee will have a way more liberall, 
Then changing hearts, to joyne them, so wee shall 
   Be one, and one anothers All.\\
\end{poem}

\begin{poem}
{The Extasie}
{John Donne}
 
Where,like a pillow on a bed, 
   A Pregnant banke swel'd up, to rest 
The violets reclining head, 
   Sat we two, one anothers best; 
Our hands were firmely cimented 
With a fast balme, which thence did spring, 
Our eye-beames twisted, and did thred 
   Our eyes, upon one double string, 
So to'entergraft our hands, as yet 
   Was all the meanes to make us one, 
And pictures in our eyes to get 
   Was all our propagation. 
As'twixt two equall Armies, Fate 
   Suspends uncertaine victorie, 
Our soules, (which to advance their state, 
   Were gone out,) hung 'twixt her, and mee. 
And whil'st our soules negotiate there, 
   Wee like sepulchrall statues lay, 
All day, the same our postures were, 
   And wee said nothing, all the day. 
If any, so by love refin'd, 
   That he soules language understood, 
And by good love were growen all minde, 
   Within convenient distance stood, 
He (though he knowes not which soule spake, 
   Because both meant, both spake the same) 
Might thence a new concoction take, 
   And part farre purer then he came. 
This Extasie doth unperplex 
   (We said) and tell us what we love, 
Wee see by this, it was not sexe 
   Wee see, we saw not what did move: 
But as all severall soules containe 
   Mixture of things, they know not what, 
Love, these mixt soules, doth mixe againe, 
   And makes both one, each this and that. 
A single violet transplant, 
   The strength, the colour, and the size, 
(All which before was poore, and scant,) 
   Redoubles still, and multiplies. 
When love, with one another so 
   Interanimates two soules, 
That abler soule, which thence doth flow, 
   Defects of lonelinesse controules. 
Wee then, who are this new soule, know, 
   Of what we are compos'd, and made, 
For, th'Atomies of which we grow, 
   Are soules, whom no change can invade. 
But O alas, so long, so farre 
   Our bodies why doe wee forbeare? 
They are ours, though not wee, Wee are 
   The intelligences, they the spheares. 
We owe them thankes, because they thus, 
   Did us, to us, at first convay, 
Yeelded their senses force to us, 
   Nor are drosse to us, but allay. 
On man heavens influence workes not so, 
   But that it first imprints the ayre, 
For soule into the soule may flow, 
   Though it to body first repaire. 
As our blood labours to beget 
   Spirits, as like soules as it can, 
Because such fingers need to knit 
   That subtile knot, which makes us man: 
So must pure lovers soules descend 
   T'affections, and to faculties, 
Which sense may reach and apprehend, 
   Else a great Prince in prison lies, 
To'our bodies turne wee then, that so 
   Weake men on love reveal'd may looke; 
Loves mysteries in soules doe grow, 
   But yet the body is his booke. 
And if some lover, such as wee, 
   Have heard this dialogue of one, 
Let him still marke us, he shall see 
   Small change, when we'are to bodies gone.\\
\end{poem}

\begin{poem}
{Love's Deitie}
{John Donne}
 
I long to talke with some old lovers ghost, 
   Who dyed before the god of Love was borne: 
I cannot thinke that hee, who then lov'd most, 
   Sunke so low, as to love one which did scorne. 
But since this god produc'd a destinie, 
And that vice-nature, custome, lets it be; 
   I must love her, that loves not mee.\\

Sure, they which made him god, meant not so much: 
   Nor he, in his young godhead practis'd it. 
But when an even flame two hearts did touch, 
   His office was indulgently to fit 
Actives to passives. Correspondencie 
Only his subject was; It cannot bee 
   Love, till I love her, that loves mee.\\

But every moderne god will now extend 
   His vast prerogative, as far as Jove. 
To rage, to lust, to write to, to commend, 
   All is the purlewe of the God of Love. 
Oh were wee wak'ned by this Tyrannie 
To ungod this child againe, it could not beo 
   I should love her, who loves not mee.\\

Rebell and Atheist too, why murmure I, 
   As though I felt the worst that love could doe? 
Love may make me leave loving, or might trie 
   A deeper plague, to make her love mee too, 
Which since she loves before, I'am loth to see; 
Falshood is worse then hate; and that must bee, 
   If shee whom I love, should love mee.\\
\end{poem}

\begin{poem}
{Love's Alchemy}
{John Donne}
 
Some that have deeper digg'd love's mine than I, 
Say, where his centric happiness doth lie; 
     I have lov'd, and got, and told, 
But should I love, get, tell, till I were old, 
I should not find that hidden mystery. 
     Oh, 'tis imposture all! 
And as no chemic yet th'elixir got, 
     But glorifies his pregnant pot 
     If by the way to him befall 
Some odoriferous thing, or medicinal, 
     So, lovers dream a rich and long delight, 
     But get a winter-seeming summer's night.\\

Our ease, our thrift, our honour, and our day, 
Shall we for this vain bubble's shadow pay? 
     Ends love in this, that my man 
Can be as happy'as I can, if he can 
Endure the short scorn of a bridegroom's play? 
     That loving wretch that swears 
'Tis not the bodies marry, but the minds, 
     Which he in her angelic finds, 
     Would swear as justly that he hears, 
In that day's rude hoarse minstrelsy, the spheres. 
     Hope not for mind in women; at their best 
     Sweetness and wit, they'are but mummy, possess'd.\\
\end{poem}

