\begin{poem}
{Morning XXVII}
{Pablo Neruda}
 
 Naked, you are simple as one of your hands,
 smooth, earthy, small, transparent, round:
 you have moon-lines, apple-pathways:
 naked, you are slender as a naked grain of wheat.\\
 
 Naked, you are blue as a night in Cuba;
 you have vines and stars in your hair;
 naked you are spacious and yellow
 as summer in a golden church.\\
 
 Naked, you are tiny as one of your nails ---
 curved, subtle, rosy, till the day is born
 and you withdraw to the underground world,\\
 
 as if down a long tunnel of clothing and of chores:
 your clear light dims, gets dressed --- drops its leaves ---
 and becomes a naked hand again.
 \end{poem}

\begin{poem}
{Lithe girl, brown girl}
{Pablo Neruda, trans. Christopher Logue}

Lithe girl, brown girl,\\

The sun that makes apples,
And stiffens the wheat,
And splits the thong weed,\\
Made your body with joy.\\
Your tongue like a red bird
Dancing on ivory,
Your lips with the smile of water.\\
You stretch your arms\\
And the sun grabs
At the loose black coils
Of your hair\\
As if water were falling.\\
Tantalize the sun, if you dare,
It will leave
Shadows that match you everywhere.\\
Lithe girl, brown girl,\\
Nothing draws me towards you,
And the heat within you
Beats me home\\
Like the sun at high noon.\\
Knowing these things,
Perhaps through knowing these things,
I seek you out,
Listening for your voice or the brush\\
Of your arms against wheat,
Or your step
Among poppies grown under water.
\end{poem}


\begin{poem}
{Clenched Soul}
{Pablo Neruda}

 We have lost even this twilight.
 No one saw us this evening hand in hand
 while the blue night dropped on the world.\\

 I have seen from my window
 the fiesta of sunset in the distant mountain tops.\\

 Sometimes a piece of sun 
 burned like a coin in my hand.\\

 I remembered you with my soul clenched
 in that sadness of mine that you know.\\

 Where were you then?
 Who else was there?
 Saying what?
 Why will the whole of love come on me suddenly
 when I am sad and feel you are far away?\\

 The book fell that always closed at twilight
 and my blue sweater rolled like a hurt dog at my feet.\\

 Always, always you recede through the evenings
 toward the twilight erasing statues.
\end{poem}

\begin{poem}
{Love Sonnet XI}
{Pablo Neruda}

 I crave your mouth, your voice, your hair.
 Silent and starving, I prowl through the streets.
 Bread does not nourish me, dawn disrupts me, all day
 I hunt for the liquid measure of your steps.\\
 
 I hunger for your sleek laugh,
 your hands the color of a savage harvest,
 hunger for the pale stones of your fingernails,
 I want to eat your skin like a whole almond.\\
 
 I want to eat the sunbeam flaring in your lovely body,
 the sovereign nose of your arrogant face,
 I want to eat the fleeting shade of your lashes,\\
 
 and I pace around hungry, sniffing the twilight,
 hunting for you, for your hot heart,
 like a puma in the barrens of Quitratue.
\end{poem}
