\begin{poem}
{The Stare's Nest by My Window}
{W. B. Yeats}

 The bees build in the crevices
 Of loosening masonry, and there
 The mother birds bring grubs and flies.
 My wall is loosening; honey-bees,
 Come build in the empty house of the stare.\\
 
 We are closed in, and the key is turned
 On our uncertainty; somewhere
 A man is killed, or a house burned.
 Yet no clear fact to be discerned:
 Come build in the empty house of the stare.\\
 
 A barricade of stone or of wood;
 Some fourteen days of civil war:
 Last night they trundled down the road
 That dead young soldier in his blood:
 Come build in the empty house of the stare.\\
 
 We had fed the heart on fantasies,
 The heart's grown brutal from the fare,
 More substance in our enmities
 Than in our love; O honey-bees,
 Come build in the empty house of the stare.\\

           ("Meditations in Time of Civil War - VI", 1928)
 \end{poem}

\begin{poem}
 {A Prayer For My Daughter}
 {W.B.Yeats}

 Once more the storm is howling, and half hid
 Under this cradle-hood and coverlid
 My child sleeps on. There is no obstacle
 But Gregory's wood and one bare hill
 Whereby the haystack- and roof-levelling wind,
 Bred on the Atlantic, can be stayed;
 And for an hour I have walked and prayed
 Because of the great gloom that is in my mind.\\

 I have walked and prayed for this young child an hour
 And heard the sea-wind scream upon the tower,
 And under the arches of the bridge, and scream
 In the elms above the flooded stream;
 Imagining in excited reverie
 That the future years had come,
 Dancing to a frenzied drum,
 Out of the murderous innocence of the sea.\\

 May she be granted beauty and yet not
 Beauty to make a stranger's eye distraught,
 Or hers before a looking-glass, for such,
 Being made beautiful overmuch,
 Consider beauty a sufficient end,
 Lose natural kindness and maybe
 The heart-revealing intimacy
 That chooses right, and never find a friend.\\

 Helen being chosen found life flat and dull
 And later had much trouble from a fool,
 While that great Queen, that rose out of the spray,
 Being fatherless could have her way
 Yet chose a bandy-legg\`{e}d smith for man.
 It's certain that fine women eat
 A crazy salad with their meat
 Whereby the Horn of Plenty is undone.\\

 In courtesy I'd have her chiefly learned;
 Hearts are not had as a gift but hearts are earned
 By those that are not entirely beautiful;
 Yet many, that have played the fool
 For beauty's very self, has charm made wise,
 And many a poor man that has roved,
 Loved and thought himself beloved,
 From a glad kindness cannot take his eyes.\\

 May she become a flourishing hidden tree
 That all her thoughts may like the linnet be,
 And have no business but dispensing round
 Their magnanimities of sound,
 Nor but in merriment begin a chase,
 Nor but in merriment a quarrel.
 O may she live like some green laurel
 Rooted in one dear perpetual place.\\

 My mind, because the minds that I have loved,
 The sort of beauty that I have approved,
 Prosper but little, has dried up of late,
 Yet knows that to be choked with hate
 May well be of all evil chances chief.
 If there's no hatred in a mind
 Assault and battery of the wind
 Can never tear the linnet from the leaf.\\

 An intellectual hatred is the worst,
 So let her think opinions are accursed.
 Have I not seen the loveliest woman born
 Out of the mouth of Plenty's horn,
 Because of her opinionated mind
 Barter that horn and every good
 By quiet natures understood
 For an old bellows full of angry wind?\\

 Considering that, all hatred driven hence,
 The soul recovers radical innocence
 And learns at last that it is self-delighting,
 Self-appeasing, self-affrighting,
 And that its own sweet will is Heaven's will;
 She can, though every face should scowl
 And every windy quarter howl
 Or every bellows burst, be happy still.\\

 And may her bridegroom bring her to a house
 Where all's accustomed, ceremonious;
 For arrogance and hatred are the wares
 Peddled in the thoroughfares.
 How but in custom and in ceremony
 Are innocence and beauty born?
 Ceremony's a name for the rich horn,
 And custom for the spreading laurel tree.\\
\end{poem}

\begin{poem}
{The Second Coming}
{William Butler Yeats}

Turning and turning in the widening gyre
The falcon cannot hear the falconer;
Things fall apart; the center cannot hold;
Mere anarchy is loosed upon the world,
The blood-dimmed tide is loosed, and everywhere
The ceremony of innocence is drowned;
The best lack all conviction, while the worst
Are full of passionate intensity.\\

Surely some revelation is at hand;
Surely the Second Coming is at hand:
The Second Coming! Hardly are those words out
When a vast image out of Spiritus Mundi
Troubles my sight: somewhere in the sands of the desert
A shape with lion body and the head of a man,
A gaze blank and pitiless as the sun,
Is moving its slow thighs, while all about it
Reel shadows of the indignant desert birds.
The darkness drops again; but now I know
That twenty centuries of stony sleep
Were vexed to nightmare by a rocking cradle,
And what rough beast, its hour come round at last,
Slouches towards Bethlehem to be born?
\end{poem}

\begin{poem}{The Wild Swans at Coole}{W B Yeats}

The trees are in their autumn beauty, 
The woodland paths are dry, 
Under the October twilight the water 
Mirrors a still sky; 
Upon the brimming water among the stones 
Are nine and fifty swans. \\

The nineteenth Autumn has come upon me 
Since I first made my count; 
I saw, before I had well finished, 
All suddenly mount 
And scatter wheeling in great broken rings 
Upon their clamorous wings. \\

I have looked upon those brilliant creatures, 
And now my heart is sore. 
All's changed since I, hearing at twilight, 
The first time on this shore, 
The bell-beat of their wings above my head, 
Trod with a lighter tread. \\

Unwearied still, lover by lover, 
They paddle in the cold, 
Companionable streams or climb the air; 
Their hearts have not grown old; 
Passion or conquest, wander where they will, 
Attend upon them still. \\

But now they drift on the still water 
Mysterious, beautiful; 
Among what rushes will they build, 
By what lake's edge or pool 
Delight men's eyes, when I awake some day 
To find they have flown away?\\
\end{poem}

 

\begin{poem}{When You Are Old}{W B Yeats}

When you are old and gray and full of sleep, 
And nodding by the fire, take down this book, 
And slowly read, and dream of the soft look 
Your eyes had once, and of their shadows deep; \\

How many loved your moments of glad grace, 
And loved your beauty with love false or true, 
But one man loved the pilgrim soul in you, 
And loved the sorrows of your changing face; \\

And bending down beside the glowing bars, 
Murmur, a little sadly, how love fled 
And paced among the mountains overhead 
And hid his face among a crowd of stars. \\
\end{poem}

\begin{poem}{Sailing to Byzantium}{W B Yeats}

That is no country for old men, The young
In one another's arms, birds in the trees
-Those dying generations-at their song,
The salmon-falls, the mackerel-crowded seas,
Fish, flesh, or fowl, commend all summer long
Whatever is begotten, born, and dies.
Caught in that sensual music all neglect
Monuments of unaging intellect.\\

An aged man is but a paltry thing,
A tattered coat upon a stick, unless
Soul clap its hands and sing, and louder sing
For every tatter in its mortal dress,
Nor is there singing school but stuying 
Monuments of its own magnificence;
And therefore I have sailed the seas and come
To the holy city of Byzantium.\\

O sages standing in God's holy fire
As in the gold mosaic of a wall,
Come from the holy fire, perne in a gyre,
And be the singing-masters of my soul.
Consume my heart away; sick with desire
And fastened to a dying animal
It knows not what it is; and gather me
Into the artifice of eternity.\\

Once out of nature I shall never take
My bodily form from any natural thing,
But such a form as Grecian goldsmiths make
Of hammered gold and gold enameling
To keep a drowsy Emperor awake;
Or set upon a golden bough to sing
To lords and ladies of Byzantium
Of what is past, or passing, or to come.\\
\end{poem}

 

\begin{poem}{Easter, 1916}{W B Yeats}

I have met them at close of day 
Coming with vivid faces 
From counter or desk among grey 
Eighteenth-century houses. 
I have passed with a nod of the head 
Or polite meaningless words, 
Or have lingered awhile and said 
Polite meaningless words, 
And thought before I had done 
Of a mocking tale or a gibe 
To please a companion 
Around the fire at the club, 
Being certain that they and I 
But lived where motley is worn: 
All changed, changed utterly: 
A terrible beauty is born.\\

That woman's days were spent 
In ignorant good will, 
Her nights in argument 
Until her voice grew shrill. 
What voice more sweet than hers 
When young and beautiful, 
She rode to harriers? 
This man had kept a school 
And rode our winged horse. 
This other his helper and friend 
Was coming into his force; 
He might have won fame in the end, 
So sensitive his nature seemed, 
So daring and sweet his thought. 
This other man I had dreamed 
A drunken, vain-glorious lout. 
He had done most bitter wrong 
To some who are near my heart, 
Yet I number him in the song; 
He, too, has resigned his part 
In the casual comedy; 
He, too, has been changed in his turn, 
Transformed utterly: 
A terrible beauty is born.\\

Hearts with one purpose alone 
Through summer and winter seem 
Enchanted to a stone 
To trouble the living stream. 
The horse that comes from the road. 
The rider, the birds that range 
From cloud to tumbling cloud, 
Minute by minute change; 
A shadow of cloud on the stream 
Changes minute by minute; 
A horse-hoof slides on the brim, 
And a horse plashes within it 
Where long-legged moor-hens dive, 
And hens to moor-cocks call. 
Minute by minute they live: 
The stone's in the midst of all.\\

Too long a sacrifice 
Can make a stone of the heart. 
O when may it suffice? 
That is heaven's part, our part 
To murmur name upon name, 
As a mother names her child 
When sleep at last has come 
On limbs that had run wild. 
What is it but nightfall? 
No, no, not night but death; 
Was it needless death after all? 
For England may keep faith 
For all that is done and said. 
We know their dream; enough 
To know they dreamed and are dead. 
And what if excess of love 
Bewildered them till they died? 
I write it out in a verse -- 
MacDonagh and MacBride 
And Connolly and Pearse 
Now and in time to be, 
Wherever green is worn, 
Are changed, changed utterly: 
A terrible beauty is born. \\
\end{poem}


\begin{poem}{The Song of Wandering Aengus}{W B Yeats}

I went out to the hazel wood, 
Because a fire was in my head, 
And cut and peeled a hazel wand, 
And hooked a berry to a thread; 
And when white moths were on the wing, 
And moth-like stars were flickering out, 
I dropped the berry in a stream 
And caught a little silver trout.\\

When I had laid it on the floor 
I went to blow the fire aflame, 
But something rustled on the floor, 
And some one called me by my name: 
It had become a glimmering girl 
With apple blossom in her hair 
Who called me by my name and ran 
And faded through the brightening air.\\

Though I am old with wandering 
Through hollow lands and hilly lands, 
I will find out where she has gone, 
And kiss her lips and take her hands; 
And wak among long dappled grass, 
And pluck till time and times are done 
The silver apples of the moon, 
The golden apples of the sun. \\
\end{poem}

 

\begin{poem}{The Lake Isle of Innisfree}{W B Yeats}

I will arise and go now, and go to Innisfree, 
And a small cabin build there, of clay and wattles made: 
Nine bean-rows will I have there, a hive for the honeybee, 
And live alone in the bee-loud glade. \\

And I shall have some peace there, for peace comes dropping slow 
Dropping from the veils of the morning to where the cricket sings; 
There midnight's all a glimmer, and noon a purple glow, 
And evenings full of the linnet's wings. \\

I will arise and go now, for always night and day 
I hear the lake water lapping with low sounds by the shore; 
While I stand on the roadway, or on the pavements gray, 
I hear it in the deep heart's core. \\
\end{poem}

 

\begin{poem}{The Scholars}{W B Yeats}

Bald heads forgetful of their sins,
Old, learned, respectable bald heads
Edit and annotate the lines
That young men, tossing on their beds,
Rhymed out in love's despair
To flatter beauty's ignorant ear.\\

All shuffle there; all cough in ink;
All wear the carpet with their shoes;
All think what other people think;
All know the man their neighbour knows.
Lord, what would they say
Did their Catullus walk that way?\\
\end{poem}

 

\begin{poem}{The Hero, the Girl, and the Fool}{W B Yeats}

The Girl. I rage at my own image in the glass,
That's so unlike myself that when you praise it
It is as though you praised another, or even
Mocked me with praise of my mere opposite;
And when I wake towards morn I dread myself
For the heart cries that what deception wins
Cruelty must keep; therefore be warned and go
If you have seen that image and not the woman.\\

The Hero. I have raged at my own strength because you have loved it.\\

The Girl. If you are no more strength than I am beauty
I had better find a convent and turn nun;
A nun at least has all men's reverence
And needs no cruelty.\\

The Hero.                  I have heard one say
That men have reverence for their holiness
And not themselves.\\

The Girl.                    Say on and say
That only God has loved us for ourselves,
But what care I that long for a man's love?\\

The Fool by the Roadside. When my days that have
From cradle run to grave
From grave to cradle run instead;
When thoughts that a fool
Has wound upon a spool
Are but loose thread, are but loose thread;\\

When cradle and spool are past
And I mere shade at last
Coagulate of stuff
Transparent like the wind,
I think that I may find
A faithful love, a faithful love.\\
\end{poem}

 

\begin{poem}{Leda and the Swan}{W B Yeats}

A sudden blow: the great wings beating still
Above the staggering girl, her thighs caressed
By the dark webs, her nape caught in his bill,
He holds her helpless breast upon his breast.\\

How can those terrified vague fingers push
The feathered glory from her loosening thighs?
And how can body, laid in that white rush,
But feel the strange heart beating where it lies?\\

A shudder in the loins engenders there
The broken wall, the burning roof and tower
And Agamemnon dead.
                               Being so caught up,
So mastered by the brute blood of the air,
Did she put on his knowledge with his power
Before the indifferent beak could let her drop?\\
\end{poem}

 

\begin{poem}{An Irish Airman Forsees His Death}{W B Yeats}

I know that I shall meet my fate 
Somewhere among the clouds above; 
Those that I fight I do not hate, 
Those that I guard I do not love; 
My country is Kiltartan Cross, 
My countrymen Kiltartan's poor, 
No likely end could bring them loss 
Or leave them happier than before. 
Nor law, nor duty bade me fight, 
Nor public men, nor cheering crowds, 
A lonely impulse of delight 
Drove to this tummult in the clouds; 
I balanced all, brought all to mind, 
The years to come seemed waste of breath, 
A waste of breath the years behind 
In balance with this life, this death. \\
\end{poem}

 

\begin{poem}{Love Song }{W B Yeats}

From the Gaelic\\

My love, we will go, we will go, I and you, 
And away in the woods we will scatter the dew; 
And the salmon behold, and the ousel too, 
My love, we will hear, I and you, we will hear, 
The calling afar of the doe and the deer. 
And the bird in the branches will cry for us clear, 
And the cuckoo unseen in his festival mood; 
And death, oh my fair one, will never come near 
In the bosom afar of the fragrant wood. \\
\end{poem}


 

\begin{poem}{Among School Children}  {W B Yeats}
  
                      \Stanza{I}
 
 I walk through the long schoolroom questioning;
 A kind old nun in a white hood replies;
 The children learn to cipher and to sing,
 To study reading-books and histories,
 To cut and sew, be neat in everything
 In the best modern way -- the children's eyes
 In momentary wonder stare upon
 A sixty-year-old smiling public man.\\
 
                     \Stanza{II}
 
 I dream of a Ledaean body, bent
 Above a sinking fire,  a tale that she
 Told of a harsh reproof, or trivial event
 That changed some childish day to tragedy --
 Told, and it seemed that our two natures blent
 Into a sphere from youthful sympathy,
 Or else, to alter Plato's parable,
 Into the yolk and white of the one shell.\\
 
                     \Stanza{III}
 
 And thinking of that fit of grief or rage
 I look upon one child or t'other there
 And wonder if she stood so at that age -- 
 For even daughters of the swan can share
 Something of every paddler's heritage -- 
 And had that colour upon cheek or hair,
 And thereupon my heart is driven wild:
 She stands before me as a living child.\\
 
                     \Stanza{IV}
 
 Her present image floats into the mind -- 
 Did Quattrocento finger fashion it
 Hollow of cheek as though it drank the wind
 And took a mess of shadows for its meat?
 And I though never of Ledaean kind
 Had pretty plumage once -- enough of that,
 Better to smile on all that smile, and show
 There is a comfortable kind of old scarecrow.\\
 
                     \Stanza{V}
 
 What youthful mother, a shape upon her lap
 Honey of generation had betrayed,
 And that must sleep, shriek, struggle to escape
 As recollection or the drug decide,
 Would think her Son, did she but see that shape
 With sixty or more winters on its head,
 A compensation for the pang of his birth,
 Or the uncertainty of his setting forth?\\
 
                     \Stanza{VI}
 
 Plato thought nature but a spume that plays
 Upon a ghostly paradigm of things;
 Solider Aristotle played the taws
 Upon the bottom of a king of kings;
 World-famous golden-thighed Pythagoras
 Fingered upon a fiddle-stick or strings
 What a star sang and careless Muses heard:
 Old clothes upon old sticks to scare a bird.\\
 
                     \Stanza{VII}
 
 Both nuns and mothers worship images,
 But those the candles light are not as those
 That animate a mother's reveries,
 But keep a marble or a bronze repose.
 And yet they too break hearts -- O presences
 That passion, piety or affection knows,
 And that all heavenly glory symbolise --
 O self-born mockers of man's enterprise;\\
 
                     \Stanza{VIII}
 
 Labour is blossoming or dancing where
 The body is not bruised to pleasure soul.
 Nor beauty born out of its own despair,
 Nor blear-eyed wisdom out of midnight oil.
 O chestnut-tree, great-rooted blossomer,
 Are you the leaf, the blossom or the bole?
 O body swayed to music, O brightening glance,
 How can we know the dancer from the dance?\\
\end{poem}


